\chapter*{Acknowledgements}

This project was funded by Biotechnology and Biological Sciences Research Council grant BB/J014575/1.

I would like to thank the many people without whom this work would not have happened.
My supervisor Mike Sternberg always offered support and allowed me to pursue my own ideas.
The whole of the Sternberg group provided discussions and a friendly environment - particularly Ioannis for making the work in Chapter~3 so much better, Suhail for keeping the computers working, Alessia for the kind words, Lawrence and Stefans for the pub chats, and Alex and Matt for the lunches over the years.
My co-supervisor David Mann and his group welcomed me in and taught me how to do experiments - Greg in particular was a great teacher, sorry for all the things I broke.
I had an enjoyable 3 months at BenevolentAI on my industry placement and want to thank everyone there for the opportunity.
Thanks also to my progress review panel, Alfonso De Simone and Matthew Fuchter, my other co-supervisor Alan Armstrong, the Stumpf group, and the open source contributors whose software I used.

I would also like to thank my family and friends, who have been so supportive over the years.
My parents and siblings always believed in me more than I believed in myself.
I have known my school friends for 14 or more years now and they still manage to surprise me.
My uni friends have provided such friendship and good times.
Various sports clubs have provided a very direct form of therapy.
And, to Joanna, for making the last year better with a convincing \textit{p}-value.
