\chapter{CDK2}
\label{cha:cdk2}

This chapter describes computational and experimental work to test a potential allosteric site on a protein important in cell cycle regulation, cyclin-dependent kinase 2 (CDK2).
% Change this to section ref
This allosteric site was predicted by ExProSE in Chapter~\ref{cha:exprose}.
A virtual screen of small molecules was carried out against the pocket.
Chosen molecules were purchased and tested...


\section{Materials and Methods}

\subsection{Virtual screening}

ZINC is a free database of commercially-available compounds for virtual screening \cite{Sterling2015}.
The ZINC12 LeadsNow subset contains over 3 million lead-like molecules in stock at chosen suppliers, where lead-like...
These are clustered using a 90\% Tanimoto cutoff as described at \url{http://zinc.docking.org} to get a representative subset of $\sim$250,000 compounds.
These compounds are docked using AutoDock Vina \cite{Trott2010} to the pocket of interest in the ANS-bound structure, PDB ID 3PXF.
The receptor and ligands are prepared in line with the AutoDock Vina documentation.
The top 2,000 structures by affinity score from this screen were retained for further docking studies.
These molecules were docked using AutoDock Vina and DOCK \cite{Allen2015} onto four structures:
\begin{itemize}
\item ...
\item ...
\item ...
\item ...
\end{itemize}
% Move these to results
Docking to multiple structures means the compounds are scored in multiple conformations of the binding site (ensemble docking).
This is important for this pocket as it is a flexible pocket.
% Mention cross-docking?

Compounds were ranked by the average of the 8 scores from the above docking (4 structures, 2 docking methods).
Violations and benign functionality \cite{Daina2017}.
Compounds removed by similarity in ChemMine \cite{Backman2011}.



\section{Results}

Native kind of there (1HCL), ANS-bound there (3PXF), cyclin-bound not there (1FIN), 4EZ3 half open.
PDB - n strucs, nothing bound apart from EDO in 3QQH/4EZ7.
Does not show up using FTMap \cite{Kozakov2015}.
Carried out virtual screening - example of docking pose by Vina/DOCK.
