\chapter{Conclusion}
\label{cha:conclusion}

This thesis has explored the concept of allostery in proteins and developed methods for allosteric site prediction.
The background of allostery, a property of the protein structural ensemble, was introduced.
Two computational approaches were described.
AlloPred uses normal modes and machine learning to predict allosteric pockets on proteins.
ExProSE uses two structures of a protein to generate ensembles that span conformational space and can be perturbed to predict and explore allostery.
Allosteric prediction methods and virtual screening were used to predict modulators for a potential allosteric site on a protein kinase important for cell cycle control, CDK2.
These modulators were tested experimentally but the results were inconclusive.
Discussion, implications and further work for each of these approaches is described in the relevant chapters.

Themes of this work have included the benefits and drawbacks of NMA, MD and distance geometry methods; the separation of finding good binding sites in general with potential allosteric sites; the importance of experimental validation of computational predictions; the difficulty of comparing allosteric prediction methods; and the necessity of taking account of the protein structural ensemble when studying allostery.
Ultimately, the variety of allosteric mechanisms and frameworks for studying allostery makes prediction challenging and indicates that more experimental and computational work is required in this important area.

For many years papers have pointed to the immense potential of allostery for both understanding and drugging proteins.
Yet they regularly contain the qualification that a unified framework of allostery remains `elusive', and approved allosteric drugs remain rare more than 50 years after the first descriptions of allostery.
In order to unlock the true potential of allostery, predictive methods need to be as established and robust as those in other areas of bioinformatics.
When allosteric prediction is as effective as prediction of secondary structure or disordered regions, the power of allostery will be truly revealed.
However the recent emergence of methods such as those presented here means the future of allosteric prediction looks bright.
In an analogous way to allostery itself, it is hoped that the effects of exploring allostery will propagate to all areas of structural biology.
